News Across Languages - Cross-Lingual Document Similarity and Event Tracking

### 2015-03-24 rupnik2015.tex
Scale polished. Jan preberi in dodaj odvzemi po želji.

Content on the internet is becoming increasingly multi-lingual. A prime example is Wikipedia - in 2001 the majority of pages were written in English, while in 2015, the percentage of English articles has dropped to 14\%. Machine translation remains relatively rudimentary - allowing people to understand simple phrases on web pages, but remain inadequate for more advanced understanding of text. At the same time, online news has begun to dominate reporting of current events. In this paper we consider the intersection of these developments: how to track events which are reported about in multiple languages.

The term event is vague and ambiguous, but for the practical purposes we define it as ``any significant happening that is being reported about in the media''. Examples of events would include shooting down of the Malaysia Airlines plane over Ukraine on July 18th, 2014 and HSBC's admittance of aiding their clients in tax evasion on February 9th, 2015 (Figure \ref{fig:event2}). Events such as these are covered by many articles and the question is how to find all the articles in different languages that are describing a single event.

\begin{figure}
\centering
\includegraphics[width=0.75\textwidth]{events2.png}
\caption{\label{fig:event2} Events are represented by collections of articles about an event, in this case the Malaysian airliner which was shot down over the Ukraine.}
\end{figure}

As input we consider a stream of articles in different languages and a list of events. Our goal is to assign articles to their corresponding events. A priori we do not know the coverage of the articles, that is not all the events may be covered and we do not know that all the articles necessarily fit into one of the events. The task is divided into two parts: detecting events within each language and then linking events across languages. In this paper we address the second step.

As input we consider a high volume of articles in different languages. By using a language detector the stream is split into separate monolingual streams. Within each monolingual stream, an online clustering approach is employed, where tracked clusters correspond to our definition of events - this is based on the Event Registry system \shortcite{Leban2014W,Leban2014I}. Our main goal in this paper is to connect such clusters (representations of events) across languages, that is detect that a set of articles in language $A$ reports on the same event as a set of articles in language $B$.

Our approach to linking clusters across languages combines two ingredients: a cross-lingual document similarity measure, which can be interpreted as a language independent topic model, and semantic annotation of documents, which enables an alternative way to comparing documents.

The first approach represents a continuation of previous work \shortcite{nips,sikdd,nips2,iti} where we explored representations of documents which were valid over multiple languages.  The representations could be interpreted as multi-lingual topics, which were then used as proxies to compute cross-lingual similarities between documents. To learn the representations, we use Wikipedia as a training corpus. Significantly, we do not only consider the major or \emph{hub} languages such as English, German, French, etc. which have significant overlap in article coverage, but also smaller languages (in terms of number of Wikipedia articles) such as Slovenian and Hindi, which may have a negligible overlap in article coverage. We can then define a similarity between any two articles regardless of language,  which allows us to cluster the articles according to topic. The underlying assumption being that articles describing the same event are similar  and will therefore be put into the same cluster.

Based on the similarity function we propose a novel algorithm for linking events/clusters across languages. The approach is based on learning a classification model from labelled data based on several sets of features. In addition to these features, cross-lingual similarity is also used to quickly identify a small list of potential linking candidates for each cluster.
This greatly increases the scalability of the system.

%As input we consider a stream of articles in different languages and a list of events. Our goal is to assign articles to their corresponding events. A priori we do not know the coverage of the articles, that is not all the events may be covered and we do not know that all the articles necessarily fit into one of the events.  To achieve our goal, we divide the task into two parts: clustering the articles according to topic (regardless of language) and matching the clusters to events.
%
%To cluster the articles, rather than rely on machine translations, we adopt well-established monolingual text mining tools to the multilingual setting. This represents the continuation of previous work \shortcite{nips,sikdd,nips2,iti} where we explored representations of documents which were valid over multiple languages.  The representations could be interpreted as multi-lingual topics, which were then used as proxies to compute cross-lingual similarities between documents. To learn the representations, we use Wikipedia as a training corpus. Significantly, we do not only consider the major or \emph{hub} languages such as English, German, French, etc. which have significant overlap in article coverage, but also smaller languages (in terms of number of Wikipedia articles) such as Slovenian and Hindi, which may have a negligible overlap in article coverage. We can then define a similarity between any two articles regardless of language,  which allows us to cluster the articles according to topic. The underlying assumption being that articles describing the same event are similar  and will therefore be put into the same cluster.
%
%The second step is to match our computed clusters to events. Since our similarity measure and clustering
%may not be perfect, we need to find the ``best" correspondence between clusters and events. In this paper, we test several different algorithms for both steps for performance in an integrated system. In particular, we test our approach within an online system for detection of world events, called Event Registry \shortcite{Leban2014W,Leban2014I}. Event Registry is a repository of events, where events are automatically identified by analyzing news articles.

%
%Event Registry works by collecting and analyzing news articles that are published by news outlets in different languages all over the world. Collected articles are first semantically annotated by identifying in them mentions of relevant concepts -- either entities or important keywords. The disambiguation and entity linking of the concepts is done using Wikipedia as the main knowledge base.
%   The key step in Event Registry is then to apply an online clustering algorithm on the articles in order to identify groups of articles that are discussing the same event.
%
%   Within this system, we test our multilingual similarity measures as well as our algorithms for matching clusters across languages.  The result is that we can test total performance in a multilingual, real-world setting.

The paper is organized as follows: we first provide an overview of the system as a whole (Section ~\ref{sec:pipeline}), followed by details on each of the two steps described above. We first introduce several approaches for measuring cross-lingual similarity (Section~\ref{sec:crosslingual}) and then the algorithms for matching clusters in different languages (Section~\ref{sec:linking}).  To improve readability, we describe relevant related work for each module in its corresponding Section. Finally, we present and interpret the experimental results, give conclusions and describe promising future directions.

\section{Pipeline}\label{sec:pipeline}

\begin{figure}[tb]
\centering
\includegraphics[width=0.75\textwidth]{pipeline1}
\caption{\label{fig:pipeline} Our pipeline - from the Event Registry we obtain a stream of multi-lingual documents and monolingual clusters. Our contribution is shown on the right, where using cross-lingual similarity, we link the clusters across languages which correspond to the same event.}
\end{figure}

@@

Our approach is based on first defining a similarity function between articles in different languages (Section~\ref{sec:crosslingual}) and then we present several efficient approaches to linking the clusters based on features (Section~\ref{sec:linking}). The pipeline is shown in Figure~\ref{fig:pipeline} First, however we describe the pre-processing on the input articles which we then take as input.

The system is designed to handle streams and so we describe how a document which enters the stream is processed. The sequence of operations is shown in  Figure~\ref{fig:erpipeline}.  A new document is first tokenized, stop words are removed and words are stemmed.
 What remains is represented in a vector-space model and normalized using $TF-IDF$ (see Section~\ref{sec:tfidf} for the definition). Since each article is tagged with its language, cosine similiarity  is used to find the modst similar existing cluster, by comparing it to the centroid vector of each cluster. A selected threshold is used to determine if the article is not similar to any existing clusters (0.4 was used in our experiments).  The article is then assigned to the corresponding cluster, otherwise a new cluster is created, initially containing only the single article.

@@

 Further details about the system can be found in \cite{}.

\begin{figure}[tbp]
\centering
\includegraphics[width=\textwidth]{ev_pipeline}
\caption{\label{fig:erpipeline}  The Event Registry pipeline, after obtaining the raw stream of multi-lingual news articles, it first cleans and processes articles individually. It then aggregates and clusters the articles into similar events (within each language) and finally provides an interface allowing us to query for articles.}
\end{figure}

@@

\section{Cross-lingual Document Similarity}
\label{sec:crosslingual}
Document similarity is an important component in techniques from text mining and natural language processing. Many techniques use the similarity as a black box, i.e., a kernel in Support Vector Machines. Comparison of documents (or other types of text snippets) is a well studied problem \todo{some references to prove that}. In this section we define document similarity in a cross-lingual setting, where the similarity function receives documents in different languages. We conclude the section by an introduction of \todo{two datasets} which we used in this paper to learn cross-lingual similarity functions.

@@

\subsection{Problem definition}
\label{sec:tfidf}
\textbf{Document representation.}

@@

\subsection{Canonical Correlation Analysis}\label{sec:CCA}
 The final approach is a statistical technique to analyze data from two sources - after which we will describe our main contribution to computing cross-lingual similarities.

@@

The main application on which we test the above similarity is cross-lingual event linking. In online media streams -- particularly new articles -- there is often duplication of reporting, different viewpoints or opinions, all centering around a single event. Same events are covered by many articles and the question we address is how to find all the articles in different languages that are describing a single event. We base our evaluation on an online system for detection of world events, called Event Registry. We do not address the problem of detection of events in this paper but rather consider the problem that given a series of events, how may we best ``tag" or match articles in different languages to these events. The events are represented by clusters of articles and so ultimately our problem reduces to find suitable matchings between clusters with articles in different languages.

@@

The problem of cross-lingual event linking is to identify the clusters describing the same events. Each article $a_i$ is written in a language $l$, where $\ell \in L = \{\ell_1,\ell_2,...,\ell_k\}$. For each language $\ell$, we generate and maintain a set of language-specific clusters $C_l$. That is, all articles are in language $l$. For any cluster $c_i \in C_{\ell_a}$, we would like to identify clusters $c_j \in C_{\ell_b}$ (where $\ell_a \neq \ell_b$) that describe the same event as $c_i$. We define \emph{equivalent clusters} as  pairs $(c_i,c_j)$ such that both clusters in the pair describe the same event in languages $C_{\ell_a}$ and $C_{\ell_b}$ respectively. Furthermore, we iterate over all the languages we consider. That is, given a cluster in language $\ell_i$, we look for matching clusters in language $\ell_j$ for $i\neq j$  in a pairwise fashion.

\begin{figure}[tb]
\centering
\includegraphics[width=0.6\textwidth]{clusters}
\caption{\label{fig:clusters}  Given clusters of articles in different languages, the goal of  linking is to match clusters which have similar content across languages. The challenge in linking is to overcome the errors in the clustering and differing content as well as make the approach scalable. Therefore, we first identify candidate clusters. }
\end{figure}

Furthermore, the matching of clusters is a \emph{generalized matching}. We cannot assume that there is only one cluster per language per event, nor can we assume complete coverage -- that there exists at least one cluster per event in every language. This implies that we cannot make any assumptions on the matching, e.g. one-to-one, complete, etc. This excludes the use of standard weighted bipartite matching type of algortihms for this problem. An example is shown in Figure~\ref{fig:clusters}, where a cluster may contain articles which are closely matched with many clusters in a different language.

We impose an additional requirement.  Identification of the equivalent clusters cannot be done exhaustively -- for a cluster $c_i$ we wish to avoid testing all other clusters for all the other languages. Performing exhaustive comparison for each cluster $c_i$ would result in $O(N^2)$ tests, where $N$ is the number of all clusters (over all languages), which is not feasible when the number of clusters is on the order of tens of thousands.

@@

Although there are a number of services that aggregate news by identifying clusters of similar articles, there are almost no services that provide linking of clusters over different languages. Google News as well as Yahoo! News are able to identify clusters of articles about same event, but they offer no linking of clusters across languages. The only service that we found, which provides cross-lingual cluster linking, is the European Media Monitor (EMM)~\cite{Steinberger2008}. EMM clusters articles in 60 languages and then tries to determine which clusters of articles in different languages describe the same event. To achieve cluster linking, EMM uses three different language independent vector representations for each cluster. First vector contains the weighted list of references to countries mentioned in the articles, while the second vector contains the weighted list of mentioned people and organizations. The last vector contains the weighted list of Eurovoc subject domain descriptors. These descriptors are topics, such as \emph{air transport}, \emph{EC agreement}, \emph{competition} and \emph{pollution control} into which articles are automatically categorized~\shortcite{Pouliquen2006}. Similarity between clusters is then computed using a linear combination of the cosine similarities computed on the lists. If the similarity is above the threshold, the clusters are linked. Compared to EMM, our approach uses document similarities to obtain a small set of potentially equivalent clusters. Additionally, we do not decide if two clusters are equivalent based on a hand-set threshold on a similarity value -- instead we use a classification model that uses a larger set of features related to the tested clusters.

@@

In order to identify clusters $c_j$ that are equivalent to cluster $c_i$, we have developed a two-stage algoirhtm. For a cluster $c_i$, we first efficiently identify a small set of candidate clusters and then find those clusters among the candidates, which are equivalent to $c_i$. An example is  shown in  Figure~\ref{fig:clusters}.

The details of the first step are described in Algorithm~\ref{cluster_merge_algo1}. The algorithm begins by individually inspecting each article $a_i$ in the cluster $c_i$. Using a chosen method for computing cross-lingual document similarity (see Section~\ref{sec:models}), it identifies the 10 most similar articles to $a_i$ in each language $l \in L$. For each  similar article $a_j$, we identify its corresponding  cluster $c_j$ and add it to the set of candidates. The set of candidate clusters obtained in this way is several orders of magnitude smaller than the number of all clusters. Although computed document similarities are approximate, our  assumption  is that articles in different languages describing the same event, will generally have a higher similarity than articles about different events. While this assumption does not always hold, redundancy in the data mitigates these false positives. Since we compute the 10 most similar articles for each article in $c_i$, we are likely to identify all the relevant candidates for cluster $c_i$.

@@

In order to determine how accurately we can predict cluster equivalence we have performed two experiments. In the first experiment we wanted to test how well do the individual approaches for cross-lingual article linking perform when used for linking the clusters about the same event. In the second experiment we tested how accurate is the prediction model when trained on different subsets of learning features. To evaluate the prediction accuracy for a given dataset we used 10-fold cross validation. In this technique, the dataset is partitioned into 10 folds, where 9 folds are used for training a model and the left-out fold is then used to test the accuracy of the model. The learning and testing process is repeated 10 times, each time with a different left-out fold.

In Section~\ref{sec:models} we have described three main algorithms for identifying similar articles in different languages. These algorithms were $k$-means, LSI and CCA. To test which of these algorithms performs best we performed the following test. Using each algorithm, we analyzed all articles in Event Registry to find the most similar articles in other languages. The values of the cross-lingual article linking features (\texttt{linkCount} and \texttt{avgSimScore}) were then recomputed for all annotated learning examples. A dataset with only these two features and the label was constructed and used to test how accurate prediction models can be induced using the data. The results of the experiment are shown in Table~\ref{table:linkingEvalAlgos}. The table shows for each of the algorithms the obtained classification accuracy, precision and recall. We can see that the most informative features for cluster linking are obtained using the CCA algorithm, while $k$-means performs the worse. Firstly we verified the AMRR (average of mean reciprocal ranks over all language pairs) performance of approaches by holding out $15000$ aligned test documents and using $300000$ aligned documents
 as the training set. Figure~\ref{pc:AMRR} shows AMRR score as the function of the number of feature vectors. It is clear that CCA outperforms LSI approach and K-means lags far behind.
Furthermore we inspected how number of features influences the accuracy of cluster linking. As we can see from Table~\ref{table:linkingEvalAlgos} choosing number of features vectors larger than $500$ barely effects linking performance. That is probably due to overfitting. Moreover we tested the performance of detecting large/small cluster linking, see Table~\ref{table:linkingEvalAlgosLargeSmall}.  As expected we obtained that it is harder to link small clusters and we need better retrieval. Hub CCA outperforms other methods in this sense also, probably as it has the best retrieval performance, see Figure~\ref{pic:AMRR}.

@@

In the second experiment we wanted to evaluate how relevant individual groups of features are to correctly determine cluster equivalence. For this purpose we tested the accuracies obtained using individual groups of features, as well as using different combination of groups. Since we found the CCA to perform best of the three algorithms, we recomputed $500$ feature vectors with better accuracy and used it to compute the values of the cross-lingual article linking features. Results of the evaluation are shown in Table~\ref{table:linkingEval}. We can see that using a single group of features, the highest prediction accuracy can be achieved using the concept related features. The classification accuracy in this case is 88.8\%. By additionally including also the cross-lingual article linking features, the classification accuracy rises slightly to 89.2\%. Using all three groups of features, the achieved accuracy is 89.6\%.
Based on the results we can make the following conclusions. The cross-lingual similarity algorithms provide valuable information that can be used to identify clusters that describe the same event in different languages. The computed features are however significantly less informative compared to the features computed on the annotated concepts. Nevertheless, the cross-lingual article similarity features are very important for two reasons. The first reason is that they allow us to identify for a given cluster a limited set of candidate clusters that are potentially equivalent. This is a very important feature since it reduces the search space for several orders of magnitude. The second reason for the importance of these features is that concept annotations are not available for all articles. Annotation of news articles is computationally intensive and can only be done for a subset of collected articles.

@@

\begin{table}[h]
\caption{Accuracy of story linking with different sets of features}
\label{table:linkingEval}
\begin{center}
\begin{tabular}{|c|c|c|c|}
  \hline
  \cline{1-4}
  Features & CA & Precision & Recall \\ \cline{1-4}
  CCA & 77.6\% & 0.783 & 0.811 \\ \cline{1-4}
  Concepts & 88.8\% & 0.884 & 0.891 \\ \cline{1-4}
  Misc & 65.3\% & 0.718 & 0.500 \\ \cline{1-4}
  CCA + concepts & 89.2\% & 0.891 & 0.893 \\ \cline{1-4}
  All & 89.6\% & 0.895 & 0.895 \\ \cline{1-4}
  \hline
\end{tabular}
\end{center}
\end{table}

@@

\subsection{Scaling}
\emph{I think that this section should be short.}
100 languages!
%Memory 100 GB per day
%Newsfeed volume 430k articles per day, 210k english, 36k Spanish, 27k German, 17k French, ....
%Number of similarities, 430k similarities every 200ms, currently could handle 1M articles,
%(n \over 2) 500X500 proxy matrices, 300000x500 Projector matrices, 430000x500 data
%Premultiply with proxy matrices, compare blockwise!
One of the main advantages of our approach is that scales really well. It is fast, very robust to quality of training data, easily extendable, simple to implement and with low machine requirements. Ut currently runs on two Intel Xeon E5-2667 v2, 3.30GHz processor machine with 256GB of RAM. That is enough to do similarity computation over 150 languages if needed. It currently uses Wikipedia as freely available knowledge base.

Currently we compute similarities over $24$ languages with tags: eng, spa, deu, zho, ita, fra, rus, swe, nld, tur, jpn, por, ara, fin, ron, kor, hrv, tam, hun, slv, pol, srp, cat, ukr but we support up to $100$ top Wikipedis languages. Our data stream is Newsfeed(http://newsfeed.ijs.si/) that produces $430k$ unique articles per day. Our system currently computes 400k similiarites per 200ms, that means that we compute $16*10^10$ similarities per day. We
store one day buffer for every language, for that we need 1.5GB of memory, we store documents as 500 dimensional vectors. For each article we compute top $10$ most similar ones in every language.
For all linear algebra matrix and vector operations we use high performance numerical libraries as BLAS, OPENBLAS and Intel MKL. This currently allows us to process more than million articles per day.
In our current implementation we use variation of hub approach. Our projector matrices are of size $500\times 300000,$ so every projector takes about $1.1$ GB of RAM. Moreover we need proxy matrices of size $500x500$ for every language pair. That is 0.5 GB for $24$ languages and $9.2$ GB for $100$ languages. All together we need around 135 GB of RAM for the system with 100 languages. Using proxy matrices we can project all the input documents in the common space. That enables us to do blockwise similarity computations and
additional improvement in our system efficiency. Our code can be therefore easily parallelized using matrix multiplication instead of doing more matrix - vector multiplications. This speeds up our code by factor around $4.$ In this way we obtain some caching gains and ability to use vectorization.
Our system is also easily extendable. If we want to add a language we only need to compute a projector matrix and proxy matrices with all other already available languages.

Comment on the caching gains in our implementation.¸¸Add the conclusion if the story.

>>>>

Content on the internet is becoming increasingly multi-lingual. A prime example is Wikipedia - in 2001 the majority of pages were written in English, while in 2012, the percentage of English articles has dropped to 14\%. Machine translation remains relatively rudimentary - allowing people to understand simple phrases on we pages, but remain inadequate for more advanced understanding of text. At the same time, online news has begun to dominate reporting of current events. In this paper we consider the intersection of these developments: how to track events which are reported about in multiple languages.

The term event is vague and ambiguous, but for the practical purposes we define it as ``any significant happening that is being reported about in the media''. Examples of events would include shooting down of the Malaysia Airlines plane over Ukraine on July 18th, 2014 and HSBC's admittance of aiding their clients in tax evasion on February 9th, 2015 (Figure \ref{}). Events such as these are covered by many articles and the question is how to find all the articles in different languages that are describing a single event.

As input we consider a stream of articles in different languages and a list of events. Our goal is to assign articles to their corresponding events. A priori we do not know the coverage of the articles, that is not all the events may be covered and we do not know that all the articles necessarily fit into one of the events.  To achieve our goal, we divide the task into two parts: clustering the articles according to topic (regardless of language) and matching the clusters to events.

To cluster the articles, rather than rely on machine translations, we adopt well-established monolingual text mining tools to the multilingual setting. This represents the continuation of previous work \shortcite{nips,sikdd,nips2,iti} where we explored representations of documents which were valid over multiple languages.  The representations could be interpreted as multi-lingual topics, which were then used as proxies to compute cross-lingual similarities between documents. To learn the representations, we use Wikipedia as a training corpus. Significantly, we do not only consider the major or \emph{hub} languages such as English, German, French, etc. which have significant overlap in article coverage, but also smaller languages (in terms of number of Wikipedia articles) such as Slovenian and Hindi, which may have a negligible overlap in article coverage. We can then define a similarity between any two articles regardless of language,  which allows us to cluster the articles according to topic. The underlying assumption being that articles describing the same event are similar  and will therefore be put into the same cluster.

The second step is to match our computed clusters to events. Since our similarity measure and clustering
may not be perfect, we need to find the ``best" correspondence between clusters and events. In this paper, we test several different algorithms for both steps for performance in an integrated system. In particular, we test our approach within an online system for detection of world events, called Event Registry \shortcite{Leban2014W,Leban2014I}. Event Registry is a repository of events, where events are automatically identified by analyzing news articles.

Event Registry works by collecting and analyzing news articles that are published by news outlets in different languages all over the world. Collected articles are first semantically annotated by identifying in them mentions of relevant concepts -- either entities or important keywords. The disambiguation and entity linking of the concepts is done using Wikipedia as the main knowledge base.
   The key step in Event Registry is then to apply an online clustering algorithm on the articles in order to identify groups of articles that are discussing the same event.

   Within this system, we test our multilingual similarity measures as well as our algorithms for matching clusters to events.  The result is that we can test total performance in a multilingual, real-world setting.

The paper is organized as follows: we first provide an overview of the system as a whole (Section ), followed by details on each of the two steps described above. We first introduce several approaches to measures of cross-lingual similarity (Section ) and then the algorithms for matching clusters to events.  To improve readability, we describe relevant related work for each module in its corresponding Section.
Finally, we present and interpret the experimental results and finally give conclusions and describe promising future directions.

\section{Pipeline}

@@

Our approach is based on first defining a similarity function between articles in different languages (Section~\ref{}) and then we present several efficient approaches to linking the clusters based on features (Section~\ref{}). First, however we describe the pre-processing on the input articles which we then take as input.

The system is designed to handle streams and so we describe how a document which enters the stream is processed.   A new document is first tokenized, stop words are removed and words are stemmed.
 What remains is represented in a vector-space model and normalized using $TF-IDF$ (see Section~\ref{} for the definition). Since each article is tagged with its language, cosine similiarity  is used to find the modst similar existing cluster, by comparing it to the centroid vector of each cluster. A selected threshold is used to determine if the article is not similar to any existing clusters (0.4 was used in our experiments).  The article is then assigned to the corresponding cluster, otherwise a new cluster is created, initially containing only the single article.

@@

 Further details about the system can be found in \cite{}, while our full pipeline is shown in Figure~\ref{}.

@@

\section{Cross-lingual Document Similarity}

Document similarity is an important component in techniques from text mining and natural language processing. Many techniques use the similarity as a black box, i.e., a kernel in Support Vector Machines. Comparison of documents (or other types of text snippets) is a well studied problem \todo{some references to prove that}. In this section we define document similarity in a cross-lingual setting, where the similarity function receives documents in different languages. We conclude the section by an introduction of \todo{two datasets} which we used in this paper to learn cross-lingual similarity functions.

@@

\textbf{Document representation.}

@@

\subsection{Canonical Correlation Analysis}
 The final approach is a statistical technique to analyze data from two sources - after which we will describe our main contribution to computing cross-lingual similarities.

@@

The main application on which we test the above similiarity is cross-lingual event linking. In online media streams -- particularly new articles -- there is often duplication of reporting, different viewpoints or opinions, all centering around a single event. Same events are covered by many articles and the question we address is how to find all the articles in different languages that are describing a single event. We base our evaluation on an online system for detection of world events, called Event Registry. We do not address the problem of detection of events in this paper but rather consider the problem that given a series of events, how may we best ``tag" or match articles in different languages to these events. The events are represented by clusters of articles and so ultimately our problem reduces to find suitable matchings between clusters with articles in different languages.

@@

The problem in the second phase of the pipeline is to identify the clusters describing the same events. Each article $a_i$ is written in a language $l$, where $\ell \in L = \{\ell_1,\ell_2,...,\ell_k\}$. For each language $\ell$, we generate and maintain a set of language-specific clusters $C_l$. That is, all articles are in language $l$. For any cluster $c_i \in C_{\ell_a}$, we would like to identify clusters $c_j \in C_{\ell_b}$ (where $\ell_a \neq \ell_b$) that describe the same event as $c_i$. We define \emph{equivalent clusters} as  pairs $(c_i,c_j)$ such that both clusters in the pair describe the same event in languages $C_{\ell_a}$ and $C_{\ell_b}$ respectively. Furthermore, we iterate over all the languages we consider. That is, given a cluster in language $\ell_i$, we look for matching clusters in language $\ell_j$ for $i\neq j$  in a pairwise fashion.

Furthermore, the matching of clusters is a \emph{generalized matching}. We cannot assume that there is only one cluster per language per event, nor can we assume complete coverage - that there exists at least one cluster per event in every language. This implies that we cannot make any assumptions on the matching, e.g. one-to-one, complete, etc. This excludes the use of standard weighted bipartite matching type of algortihms for this problem.

We impose an additional requirement.  Identification of the equivalent clusters cannot be done exhaustively -- for a cluster $c_i$ we wish to avoid testing all other clusters for all the other languages. Performing exhaustive comparison for each cluster $c_i$ would result in $O(N^2)$ tests, where $N$ is the number of all clusters (over all languages), which is not feasible for when the number of clusters is on the order of tens of thousands.

@@

Although there are a number of services that aggregate news by identifying clusters of similar articles, there are very few service that provide linking of clusters over different languages. Google News as well as Yahoo! News are able to identify clusters of articles about same event, but they offer no linking of clusters across languages. The only service that provides cross-lingual cluster linking is the European Media Monitor (EMM)~\cite{Steinberger2008}. EMM clusters articles in 60 languages and then tries to determine which clusters of articles in different languages describe the same event. To achieve cluster linking, EMM uses three different language independent vector representations for each cluster. First vector contains the weighted list of references to countries mentioned in the articles, while the second vector contains the weighted list of mentioned people and organizations. The last vector contains the weighted list of Eurovoc subject domain descriptors. These descriptors are topics, such as \emph{air transport}, \emph{EC agreement}, \emph{competition} and \emph{pollution control} into which articles are automatically categorized~\shortcite{Pouliquen2006}. Similarity between clusters is then computed using a linear combination of the cosine similarities computed on the lists.

@@

In order to identify clusters $c_j$ that are equivalent to cluster $c_i$, we have developed a two-stage algoirhtm. First, for a cluster $c_i$, we efficiently identify a small set of candidate clusters and then  find the clusters from among the candidates which are equivalent to $c_i$.

The details of the first step are described in Algorithm~\ref{cluster_merge_algo1}. The algorithm begins by individually inspecting each article $a_i$ in the cluster $c_i$. Using CCA  (see Section~\ref{}),  it then identifies the 10 most similar articles to $a_i$ in each language $l \in L$. For each  similar article $a_j$, we identify its corresponding  cluster $c_j$ and add it to the set of candidates. \todo{we should say a few words why this is more efficient - since now we are comparing each article to all other articles - while we said it is not feadible for clusters}. While the CCA simliarity is approximate, our  assumption  is that articles in different languages describing the same event, will generally have a higher similarity than articles about different events. While this assumption does not always hold, redundancy in the data mitigates these false positives. Since we compute the 10 most similar articles for each article in $c_i$, we are likely to identify all the relevant candidates for cluster $c_i$.

@@

In order to determine how accurately we can predict cluster equivalence we have performed two experiments in multi-lingual setting using English, German and Spanish language for which we had labeled data to evaluate the linking performance. In the first experiment we wanted to test how well do the individual approaches for cross-lingual article linking perform when used for linking the clusters about the same event. In the second experiment we tested how accurate is the prediction model when trained on different subsets of learning features. To evaluate the prediction accuracy for a given dataset we used 10-fold cross validation. In this technique, the dataset is partitioned into 10 folds, where 9 folds are used for training a model and the left-out fold is then used to test the accuracy of the model. The learning and testing process is repeated 10 times, each time with a different left-out fold.

In Section~\ref{sec:models} we have described three main algorithms for identifying similar articles in different languages. These algorithms were $k$-means, LSI and CCA. To test which of these algorithms performs best we performed the following test. As a training set we used common Wikipedia alignment for all three languages. Using each algorithm, we analyzed all articles in Event Registry to find the most similar articles in other languages. The values of the cross-lingual article linking features (\texttt{linkCount} and \texttt{avgSimScore}) were then recomputed for all annotated learning examples. A dataset with only these two features and the label was constructed and used to test how accurate prediction models can be induced using the data. The results of the experiment are shown in Table~\ref{table:linkingEvalAlgos}. The table shows for each of the algorithms the obtained classification accuracy, precision and recall. We can see that the most informative features for cluster linking are obtained using the CCA or the LSI algorithm, while $k$-means performs the worse. To decide which method to use we made some further experiments. Firstly we verified the AMRR (average of mean reciprocal ranks over all language pairs) performance of approaches on Wikipedia data by holding out $15000$ aligned test documents and using $300000$ aligned documents
 as the training set. Figure~\ref{pic:AMRR} shows AMRR score as the function of the number of feature vectors. It is clear that CCA outperforms LSI approach and K-means lags far behind when testing on Wikipedia data. Similar performance of the LSI and CCA for cluster linking can be explained by domain missmatch, we train and test our model  AMRR performance on Wikipedia data but at the end we apply it to the news domain. Sadly we do not have any readily available data for this domain so we were not able to test properly.
Furthermore we inspect how number of features influences the accuracy of cluster linking. As we can see from Table~\ref{table:linkingEvalAlgos} choosing number of features vectors larger than $500$ barely effects linking performance. That is probably due to overfitting. Moreover we tested the performance of detecting large/small cluster linking, see Table~\ref{table:linkingEvalAlgosLargeSmall}. Large clusters aggregate a lot of information so the accuracy is not so important as long as we assign to correct clusters. In the case of small clusters accuracy should matter more.  As expected we obtained that it is harder to link small clusters and we need better retrieval accuracy in such cases. CCA outperforms other methods in this sense, probably as it has the best retrieval performance, see Figure~\ref{pic:AMRR}. Added accuracy is also beneficial when we want
to pick good representatives of clusters. Therefore we decided to use CCA for cluster linking.

@@

In the second experiment we wanted to evaluate how relevant individual groups of features are to correctly determine cluster equivalence. For this purpose we tested the accuracies obtained using individual groups of features, as well as using different combination of groups. Since we found the CCA to perform best of the three algorithms, we used it to compute the values of the cross-lingual article linking features. Results of the evaluation are shown in Table~\ref{table:linkingEval}. We can see that using a single group of features, the highest prediction accuracy can be achieved using the concept related features. The classification accuracy in this case is 88.8\%. By additionally including also the cross-lingual article linking features, the classification accuracy rises slightly to 89.2\%. Using all three groups of features, the achieved accuracy is 89.6\%.
Based on the results we can make the following conclusions. The cross-lingual similarity algorithms provide valuable information that can be used to identify clusters that describe the same event in different languages. The computed features are however significantly less informative compared to the features computed on the annotated concepts. Nevertheless, the cross-lingual article similarity features are very important for two reasons. The first reason is that they allow us to identify for a given cluster a limited set of candidate clusters that are potentially equivalent. This is a very important feature since it reduces the search space for several orders of magnitude. The second reason for the importance of these features is that concept annotations are not available for all articles. Annotation of news articles is computationally intensive and can only be done for a subset of collected articles.

@@

\begin{table}[h]
\caption{Accuracy of story linking with different sets of features}
\label{table:linkingEval}
\begin{center}
\begin{tabular}{|c|c|c|c|}
  \hline
  \cline{1-4}
  Features & CA & Precision & Recall \\ \cline{1-4}
  CCA & 78.2\% & 0.763 & 0.816 \\ \cline{1-4}
  Concepts & 88.8\% & 0.884 & 0.891 \\ \cline{1-4}
  Misc & 65.3\% & 0.718 & 0.500 \\ \cline{1-4}
  CCA + concepts & 89.2\% & 0.891 & 0.893 \\ \cline{1-4}
  All & 89.6\% & 0.895 & 0.895 \\ \cline{1-4}
  \hline
\end{tabular}
\end{center}
\end{table}

@@

\subsection{Scaling}
\emph{I think that this section should be short.}
100 languages!
%Memory 100 GB per day
%Newsfeed volume 430k articles per day, 210k english, 36k Spanish, 27k German, 17k French, ....
%Number of similarities, 430k similarities every 200ms, currently could handle 1M articles,
%(n \over 2) 500X500 proxy matrices, 300000x500 Projector matrices, 430000x500 data
%Premultiply with proxy matrices, compare blockwise!
One of the main advantages of our approach is that scales really well. It is fast, very robust to quality of training data, easily extendable, simple to implement and with low machine requirements. The similarity pipeline is the most demanding part and currently runs on two Intel Xeon E5-2667 v2, 3.30GHz processor machine with 256GB of RAM. That is enough to do similarity computation over 150 languages if needed. It currently uses Wikipedia as freely available knowledge base.
Similarity pipeline dramastically reduces the search space when linking clusters.

Currently we compute similarities over $24$ languages with tags: eng, spa, deu, zho, ita, fra, rus, swe, nld, tur, jpn, por, ara, fin, ron, kor, hrv, tam, hun, slv, pol, srp, cat, ukr but we support up to $100$ top Wikipedis languages. Our data stream is Newsfeed (http://newsfeed.ijs.si/) that produces $430k$ unique articles per day. Our system currently computes 400k similiarites per 200ms, that means that we compute $16 \cdot 10^{10}$ similarities per day. We
store one day buffer for every language, for that we need 1.5 GB of memory, we store documents as 500 dimensional vectors. We must note that time complexity of similarity computations scales linearly with dimension of the feature space and does not  depend of number of languages. For each article we compute top $10$  most similar ones in every other language.
For all linear algebra matrix and vector operations we use high performance numerical linear algebra libraries as BLAS, OPENBLAS and Intel MKL. This currently allows us to process more than million articles per day.
In our current implementation we use variation of hub approach. Our projector matrices are of size $500\times 300000,$ so every projector takes about $1.1$ GB of RAM. Moreover we need proxy matrices of size $500\times500$ for every language pair. That is 0.5 GB for $24$ languages and $9.2$ GB for $100$ languages. All together we need around 135 GB of RAM for the system with 100 languages.
 Usage of proxy matrices enables projection of all input documents in the common space and handling language pairs with missing or low alignment. That enables us to do block-wise similarity computations and
additional improvement in our system efficiency. Our code can be therefore easily parallelized using matrix multiplication instead of doing more matrix - vector multiplications. This speeds up our code by factor around $4.$ In this way we obtain some caching gains and ability to use vectorization.
Our system is also easily extendable. If we want to add new language we only need to compute a projector matrix and proxy matrices with all other already available languages.

Add the conclusion of the story and the wrap up, Andrej needs help with this.
We presented the system ...
